---
title: "Untitled"
format: html
editor: visual
---

## Function

### About

-   A set of statements which perform a specific task
-   R has a large number of in-built functions

### Syntax

``` r
function_name <- function(arg_1, arg_2) { 
  function_body 
  }
```

### Types

| Types        | Description                     |
|--------------|---------------------------------|
| Built-in     | Pre-defined function            |
| User-defined | Defined by user using R program |

### Built-in

| Function | Description  |
|----------|--------------|
| sum()    | Summation    |
| mean()   | Mean value   |
| median() | Median value |
| min()    | Minimum      |
| max()    | Maximum      |

### Example 1

```{r}
x <- 1:100

sum(x)
mean(x)
median(x)
min(x)
max(x)
```

### Special

| Function | Description |
|----|----|
| c() | Combines values into a vector. |
| print() | Displays output in the console. Auto-print is disabled in loops/function |
| paste() | Concatenates strings (or variables) into single string. |
| cat() | Prints output without quotes/indices. Useful for clean logging. |

### Examples

```{r}
vector <- c(2, 5, -7, -8, 10)
```

```{r}
print(vector)
```

```{r}
paste("Vector", vector)
cat("Vector", vector)
```

# Control Statement

## For Loop

### About

-   Loop repeats a specific block of code.
-   A for loop is used to iterate (repeat) over a sequence.

### Syntax

``` r
for (val in sequence)
{
statement
}
```

### Example

-   Count the number of even numbers in a sequence.

```{r}
for (i in vector) {
  i^2
}
```

```{r}
for (i in vector) {
 	print(i^2)
}
```

## Your Turn 01

### Question

-   Write a for loop in R that iterates over each value in a vector.
-   If value greater (\>) than 0, print the value to the console."

### Answer

```{r}







```

## while loop

### Syntax

``` r
while (test_expression)
{
statement
}
```

### Example

```{r}
i <- 1
while (i < 6) {
  print(i)
  i = i+1
}
```

# Break and Next Statement

## Break

### Syntax

``` r
if (test_expression) {
break
}
```

### Example 1

```{r}
x <- 1:5
for (i in x) {
  if (i == 4){
  break
  }
  print(i)
}
```

## Next

### Syntax

``` r
if (test_condition) {
next
}
```

### Example

```{r}
x <- 1:5
for (i in x) {
  if (i == 4){
  next
  }
  print(i)
}
```

## Repeat Loop

### Syntax

``` r
repeat {
statement
}
```

### Example

```{r}
x <- 1

repeat {
  print(x)
  x = x+1
  if (x == 5){
  break
  }
}
```

# Function II

### Types

| Types        | Description                     |
|--------------|---------------------------------|
| Built-in     | Pre-defined function            |
| User-defined | Defined by user using R program |

## User-defined Function

### Example

```{r}
powr <- function(x, y) {

  output <- x^y
  paste(x,"^", y, "is", output)
}


powr(4, 5)
```

### Named arguments

Named argument will not affect positional arguments.

```{r}
powr(x=4, y=5)
powr(y=5, x=2)
```

### Default value

Default value for arguments.

```{r}
powr <- function(x, y=2) {

  output <- x^y
  print(paste(x,"^", y, "is", output))
}


powr(4)

powr(4,5)
```

## Return a value

### Syntax

return(expression)

### No Return

Function without return

```{r}
check <- function(x) {
  if (x > 0) {
  result <- "Positive"
  }
  else if (x < 0) {
  result <- "Negative"
  }
  else {
  result <- "Zero"
  }
}

check(-10)

```

### Return

```{r}
check <- function(x) {
if (x > 0) {
result <- "Positive"
}
else if (x < 0) {
result <- "Negative"
}
else {
result <- "Zero"
}
  
return(result)

}

check(10)
```

### Conditional return

```{r}
check <- function(x) {
  if (x>0) {
  return("Positive")
  }
  else if (x<0) {
  return("Negative")
  }
  else {
  return("Zero")
  }
}

check(0)
```

## Your Turn 02

### Question

Create a function `calculate_ticket_price()` below determines the base price of a movie ticket based on age:

-   Children (age \< 13): \$5
-   Seniors (age \> 65): \$7
-   Other (age 13–65): \$10

### Answer

```{r}











```

## Function's Scope

-   global variable is shared across the program.
-   local variable only exist in certain part of program like function.

```{r}
a <- "apple" # global variable

fun_scope <- function(){
  a <- "ball" # local variable
  return(a)
}

fun_scope() # call function
print(a) # print a variable
```
